### 2011-06-30 ivcj_lfw.tex

\documentclass[preprint, 12pt]{elsarticle}
%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

%\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
%\overrideIEEEmargins

\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{lineno}
\usepackage{color}

%===========================================================
\usepackage{multirow}
\usepackage[tight,normalsize,sf,SF]{subfigure}

\usepackage{array} \newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}
\newcommand{\foreign}[1]{\emph{#1}}

%===========================================================

%\author{Anonymous Submission}
\journal{Image and Vision Computing Journal}

\begin{document}

\begin{frontmatter}

\title{
Beyond Simple Features: A Large-Scale Feature Search Approach to Unconstrained Face Recognition
}

\author[rowland,mit]{Nicolas Pinto}
\ead{pinto@rowland.harvard.edu}

\author[rowland]{David D. Cox}
\ead{cox@rowland.harvard.edu}

\address[rowland]{The Rowland Institute at Harvard, Harvard University, Cambridge, MA 02142}
\address[mit]{McGovern Institute for Brain Science at MIT, Cambridge, MA 02139}

%===========================================================
\begin{abstract}
Many modern computer vision algorithms are built atop of a set of low-level
feature operators (such as SIFT \cite{sift,luo2007person}; HOG \cite{dalal2005hog,albiol2008face}; or LBP
\cite{ahonen2004face,ahonen2006face}) that transform raw pixel values into a representation better
suited to subsequent processing and classification.  While the choice of feature
representation is often not central to the logic of a given algorithm, the
quality of the feature representation can have critically important implications
for performance.  Here, we demonstrate a large-scale feature search approach to
generating new, more powerful feature representations in which a multitude of
complex, nonlinear, multilayer neuromorphic feature representations are randomly
generated and screened to find those best suited for the task at hand.  In
particular, we show that a brute-force search can generate representations that,
in combination with standard machine learning blending techniques, achieve
state-of-the-art performance on the \emph{Labeled Faces in the Wild (LFW)} \cite{huang:lfw}
unconstrained face recognition challenge set.  These representations outperform
previous state-of-the-art approaches, in spite of requiring less training data
and using a conceptually simpler machine learning backend.  We argue that such
large-scale-search-derived feature sets can play a synergistic role with other
computer vision approaches by providing a richer base of features with which to
work.

\end{abstract}

\begin{keyword}
face recognition \sep biologically-inspired
\end{keyword}

\end{frontmatter}

% Neuromorphic vision systems -- those that draw inspiration from the brain --
% have been demonstrated to be a powerful class of algorithms for a variety of
% face and object recognition tasks
% \cite{serre2007ror,mutch2008ocr,pinto:plos08,pinto:eccv08,jarrett-iccv-09}.

% We show that
% minor adjustment of the comparison function used can bring a simple,
% single-layer \emph{V1-like} model \cite{pinto:plos08} to within a few percent of
% state-of-the-art performance.  Further, by employing high-throughput screening
% and simple kernel blending techniques, we show that more sophisticated
% multi-layer architectures \cite{pinto:plos09} can achieve
% state-of-the-art performance.  To the extent that one accepts that the LFW set
% is a good surrogate for the problem of unconstrained face verification, these
% results show that our neuromorphic models are competitive with other
% state-of-the-art approaches, even without invoking particularly sophisticated
% machine learning techniques.  At the same time, we present an analysis of the
% errors made by our various models, and show that each of them makes appreciably
% the same errors, and that a large fraction of errors can be qualitatively
% explained by variation in the view of the targets. We argue that seriously
% tackling such image variation, and building sets that contain more real-world
% variation, will be an essential component of future research in unconstrained
% face recognition.

% -----------------------------------------------------------------------------
\section{Introduction}
% -----------------------------------------------------------------------------

\input{introduction.tex}

%------------------------------------------------------------------------------
\section{Methods}
% -----------------------------------------------------------------------------

\input{methods.tex}

% -----------------------------------------------------------------------------
\section{Results}
% -----------------------------------------------------------------------------

\input{results.tex}

% -----------------------------------------------------------------------------
\section{Discussion}
% -----------------------------------------------------------------------------

\input{discussion.tex}

% -----------------------------------------------------------------------------
\section{Acknowledgments}
% -----------------------------------------------------------------------------
The authors would like to thank Hanspeter Pfister, Wen-Mei Hwu, Volodymyr
Kindratenko and Jeremy Enos for making additional GPU clusters available for
this work. This work was funded by the Rowland Institute of Harvard, the NVIDIA
Graduate Fellowship, and the National Science Foundation (IIS 0963668).
Hardware support was generously provided by the NVIDIA Corporation.

\bibliographystyle{elsarticle-harv}
\bibliography{ivcj_lfw}

\end{document}